Staph killing assays

### 2020-07-28 Microscopy Data Clean.Rmd
updated figures

plotNonButterfly(bin5and100Hist,"Sa", "PA14wt")
plotNonButterfly(bin5and100Hist,"Sa", "PA14mut")
plotNonButterfly(bin5and100Hist,"Pa", "PA14wt")
plotNonButterfly(bin5and100Hist, "Pa", "PA14mut")
```

>>>>

plotNonButterfly(bin5and100Hist,"Sa", "PA14wt")
plotNonButterfly(bin5and100Hist,"Sa", "PA14mut")
plotNonButterfly(bin5and100Hist,"Pa", "PA14wt")
plotNonButterfly(bin5and100Hist, "Pa", "PA14mut")
```
```{r Aggregate histograms planktonic vs aggregte}

#this is the summary beacuse I dont like further analyzing data after summarizing shit, so mutate at the top follows 
hist.plank.Agg.Sumary<-hist.5 %>%  group_by(Species, Sample,Condition,Strain)%>% 
  mutate(Aggregation = case_when(max == 5 ~ "Planktonic",
                        max>5 ~ "Aggregate")) %>% ungroup() %>% 
  group_by(Species,Sample,Condition,ConditionNew, Strain,Aggregation) %>% summarize(FrequencyAggregation = sum(frequency), BiomassAggregation = sum(BiomassBin), DensityAggregation = sum(Density))
#now I have to do some t-tests, since I used the same cultures for mono and co for each replicate, paired t-tests are allowed.
##### 
#I know this sucks, but it's 1 AM so idgaf
SaMonoWtAggSum<-hist.plank.Agg.Sumary %>%filter(Condition == "Sa Mono", Strain == "PA14wt",Aggregation == "Aggregate") %>% pull(BiomassAggregation)
SaCoWtAggSum<-hist.plank.Agg.Sumary %>%filter(Species =="Sa", ConditionNew == "Co", Strain == "PA14wt",Aggregation == "Aggregate") %>% pull(BiomassAggregation)
SaWtAggTest<-t.test(x = SaMonoWtAggSum, y = SaCoWtAggSum, paired = T)
SaMonoMutAggSum<-hist.plank.Agg.Sumary %>%filter(Condition == "Sa Mono", Strain == "PA14mut",Aggregation == "Aggregate") %>% pull(BiomassAggregation)
SaCoMutAggSum<-hist.plank.Agg.Sumary %>%filter(Species == "Sa", ConditionNew == "Co", Strain == "PA14mut",Aggregation == "Aggregate") %>% pull(BiomassAggregation)
SaMutAggTest<-t.test(x = SaMonoMutAggSum, y = SaCoMutAggSum, paired = T)

PaMonoWtAggSum<-hist.plank.Agg.Sumary %>%filter(Condition == "Pa Mono", Strain == "PA14wt",Aggregation == "Aggregate") %>% pull(BiomassAggregation)
PaCoWtAggSum<-hist.plank.Agg.Sumary %>%filter(Species == "Pa", ConditionNew == "Co", Strain == "PA14wt",Aggregation == "Aggregate") %>% pull(BiomassAggregation)
PaWtAggTest<-t.test(x = PaMonoWtAggSum, y = PaCoWtAggSum, paired = T)
PaMonoMutAggSum<-hist.plank.Agg.Sumary %>%filter(Condition == "Pa Mono", Strain == "PA14mut",Aggregation == "Aggregate") %>% pull(BiomassAggregation)
PaCoMutAggSum<-hist.plank.Agg.Sumary %>%filter(Species == "Pa", ConditionNew == "Co", Strain == "PA14mut",Aggregation == "Aggregate") %>% pull(BiomassAggregation)
PaMutAggTest<-t.test(x = PaMonoMutAggSum, y = PaCoMutAggSum, paired = T)

SaMonoWtVsMutAggTest<-t.test(x = SaMonoWtAggSum, y = SaMonoMutAggSum, paired = F)
SaCoWtVsMutAggTest<-t.test(x = SaCoWtAggSum, y = SaCoMutAggSum, paired = F)
PaMonoWtVsMutAggTest<-t.test(x = PaMonoWtAggSum, y = PaMonoMutAggSum, paired = F)
PaCoWtVsMutAggTest<-t.test(x = PaCoWtAggSum, y = PaCoMutAggSum, paired = F)

tTestDataAggvsPlank<-list(SaWtAggTest, SaMutAggTest, PaWtAggTest,PaMutAggTest,SaMonoWtVsMutAggTest,SaCoWtVsMutAggTest,PaMonoWtVsMutAggTest,PaCoWtVsMutAggTest)
TtestTable<-name <- function(dataPoint) {
  data <-data.frame(dataPoint$data.name, dataPoint$p.value)
}
tTestDataTable<-bind_rows(map(tTestDataAggvsPlank, TtestTable))
#so, it turns out I only had to do either agg or plank since the variance is the same. COOL...

#####
#this is using mutate which is the same as above

hist.plank.Agg.Stats<-hist.plank.Agg.Sumary %>% group_by(Species,Condition, ConditionNew,Strain,Aggregation) %>%
  summarise(MeanFreq = mean(FrequencyAggregation), SEFreq = std.error(FrequencyAggregation), MeanBiomass = mean(BiomassAggregation), SEBiomass = std.error(BiomassAggregation))

PlotAggsPlank<- function(strain, titulo){
ggplot()+
  geom_col(data = hist.plank.Agg.Stats %>% filter(Strain == strain),aes(x=ConditionNew,y=MeanBiomass, fill = Aggregation),size=3,position = position_dodge( width = .9))+
  geom_errorbar(data = hist.plank.Agg.Stats%>% filter(Strain == strain),aes(x=ConditionNew,ymin=MeanBiomass-SEBiomass, ymax = MeanBiomass+SEBiomass),size=.5,position = position_dodge2( width = .6))+
  facet_grid(.~Species)+
  labs(title = titulo, x = "", y = "Biomass contribution")+
  theme_cowplot()+
  theme(plot.title = element_text(hjust = 0.5))
}
PlotAggsPlank("PA14wt", "wt")
PlotAggsPlank("PA14mut", "pqsL")

```
```{r Aggregate histograms logarithmic bins}
#Now I need to do histograms with bins in a log scale, but for this I want to use the raw data that includes biomass per bin and density.
#let's start by revisiting the original data set and just filter one sample to simplify things:
hist.5.rep1<-hist.5 %>% filter(Sample == "rep1"&Condition == "Pa Mono"&Strain=="PA14mut")
hist.5.rep1.logBins<-hist.5.rep1 %>% 
  mutate(LogBin = case_when(max <= 10 ~ 10,
                        max<= 100    ~ 100,
                        max<= 1000   ~ 1000,
                        max<= 10000  ~ 10000,
                        max<= 100000 ~ 100000,
                        max > 100000 ~ 1000000)) %>% 
  group_by(LogBin) %>% mutate(FrequencyLogBin = sum(frequency), BiomassLogBin = sum(BiomassBin), DensityLogBin = sum(Density))
#Now this is the real case
hist.5.LogBin<-hist.5 %>%  group_by(Species, Sample,Condition,Strain)%>% 
  mutate(LogBin = case_when(max <= 10 ~ 10,
                        max<= 100    ~ 100,
                        max<= 1000   ~ 1000,
                        max<= 10000  ~ 10000,
                        max<= 100000 ~ 100000,
                        max > 100000 ~ 1000000)) %>% ungroup() %>% 
  group_by(Species, Sample,Condition,Strain,LogBin) %>% mutate(FrequencyLogBin = sum(frequency), BiomassLogBin = sum(BiomassBin), DensityLogBin = sum(Density))

#this is easier to put in bar graphs
hist.5.LogBinSummarize<-hist.5 %>% filter(!max == 5)%>%  group_by(Sample,Species,Condition,Strain)%>% 
  mutate(LogBin = case_when(max <= 10 ~ 10,
                        max<= 100    ~ 100,
                        max<= 1000   ~ 1000,
                        max<= 10000  ~ 10000,
                        max<= 100000 ~ 100000,
                        max > 100000 ~ 1000000)) %>% ungroup() %>% 
  group_by(Species, Sample,Condition,ConditionNew,Strain,LogBin) %>% summarise(FrequencyLogBin = sum(frequency), BiomassLogBin = sum(BiomassBin), DensityLogBin = sum(Density))

PlotFreqLogBin<-function(strain, species,titulo){
  df <-hist.5.LogBinSummarize%>% filter(Strain == strain & Species == species)
  ggplot()+
    geom_col(data = df,aes(x=as.factor(LogBin),y=FrequencyLogBin,  fill =Sample ),size=4)+
    #geom_point(data = df,aes(x=as.factor(LogBin),y=FrequencyLogBin,  shape =Sample ),size=4)+
    geom_label(data = df %>% filter(Sample == "rep1"),aes(x=as.factor(LogBin),y=1, label =round(BiomassLogBin*100,digits=1), color = "rep1"),size=4)+
    geom_label(data = df%>% filter(Sample == "rep2"),aes(x=as.factor(LogBin),y=.325, label =round(BiomassLogBin[Sample=="rep2"]*100,digits=1), color = "rep2"),size=4)+
    geom_label(data = df%>% filter(Sample == "rep3"),aes(x=as.factor(LogBin),y=.1, label =round(BiomassLogBin[Sample=="rep3"]*100,digits=1), color = "rep3"),size=4)+
    labs(title = titulo)+
    scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    #scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    facet_grid(.~ConditionNew)+
    labs(y = "Frequency",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    guides(color = FALSE)+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
}
PlotFreqLogBin("PA14wt", "Sa", "Sa WT")
PlotFreqLogBin("PA14wt", "Pa", "Pa WT")
PlotFreqLogBin("PA14mut", "Sa", "Sa pqsL")
PlotFreqLogBin("PA14mut", "Pa", "Pa pqsL")

### 2020-07-28 Paper Figures/Paper figures Scipts.Rmd
updated figures

```{r code Figure 1}
GraphGrowthCurveLog1<-function (rawData, statsData){
    shakingWtPlot<-ggplot()+
      geom_line(data=statsData%>% filter(condition == "Shaking" & strain == "wt"), aes(x=time, y=logCFUmean, linetype = bug), size=.5)+
      geom_errorbar(data=statsData%>% filter(condition == "Shaking" & strain == "wt"),aes(x = time, ymin = logCFUmean-logCFUse, ymax=logCFUmean+logCFUse), width = 0, size = 0.5)+
      geom_hline(yintercept=9.619, color='gray', linetype = "solid") + #Pa stirred
      geom_hline(yintercept=9.13, color='gray', linetype = "dashed") + #Sa stirred
      scale_linetype_manual(values = c( "solid","dashed"), labels = c("P.aeruginosa","S. aureus"))+
      labs(x = "Time (hours)", y = expression(paste(Log[10], " CFU")), linetype = "Species")+
      scale_y_continuous(limits = c(4,10))+
      scale_x_continuous(breaks = c(4,8,12,16))+
      theme_cowplot(12)

    staticWtPlot<-ggplot()+
      geom_line(data=statsData%>% filter(condition == "Static" & strain == "wt"), aes(x=time, y=logCFUmean, linetype = bug), size=.5)+
      geom_errorbar(data=statsData%>% filter(condition == "Static" & strain == "wt"),aes(x = time,ymin=logCFUmean-logCFUse, ymax = logCFUmean+logCFUse), width = 0, size = 0.5)+
      scale_linetype_manual(values = c( "solid","dashed"), labels = c("stirred","static"))+
      geom_hline(yintercept=8.88, color='gray', linetype = "dashed") +
      geom_hline(yintercept=9.07, color='gray', linetype = "solid") +
      labs(x = "Time (hours)", y = expression(paste(Log[10], " CFU")), linetype = "")+
      scale_y_continuous(limits=c(4,10))+
      scale_x_continuous(breaks = c(4,8,12,16))+
      theme_cowplot(12)

    legend <- get_legend(shakingWtPlot + theme(legend.box.margin = margin(0, 0, 0, -10)))
    prow <- plot_grid(shakingWtPlot+theme(legend.position="none"),staticWtPlot+theme(legend.position="none"),
                      align='vh',labels = c("Stirred", "Static"),hjust = c(-3, -3.8),nrow = 1)
    plot_grid(prow,legend,rel_widths = c(3, .4))
    }
#Graph logs
GraphGrowthCurveLog1(PA14Data,PA14DataStats)
```
Figures 2 and 5 are on document : "Staph killing assays.Rmd"

```{r Figure 4}
GraphGrowthCurveLog2<-function (rawData, statsData){
  shakingMutPlot<-ggplot()+
    geom_line(data=statsData%>% filter(condition == "Shaking" & strain ==  "mut"), aes(x=time, y=logCFUmean, linetype = bug), size=.5)+
    geom_errorbar(data=statsData%>% filter(condition == "Shaking" & strain == "mut"), aes(x = time, ymin = logCFUmean-logCFUse, ymax = logCFUmean+logCFUse), width = 0, size = 0.5)+
    scale_linetype_manual(values = c( "solid","dashed"), labels = c("P.aeruginosa","S. aureus"))+
    geom_hline(yintercept=9.13, color='gray', linetype = "dashed") +
    geom_hline(yintercept=9.57, color='gray', linetype = "solid") +
    labs(x = "Time (hours)", y = expression(paste(Log[10], " CFU")), linetype = "Species")+
    scale_y_continuous(limits = c(4,10))+
    scale_x_continuous(breaks = c(4,8,12,16))+
    theme_cowplot(12)

  staticMutPlot<-ggplot()+
    geom_line(data=statsData%>% filter(condition == "Static" & strain == "mut"), aes(x=time, y=logCFUmean, linetype = bug), size=.5)+
    geom_errorbar(data=statsData%>% filter(condition == "Static" & strain ==  "mut"), aes(x = time, ymin = logCFUmean-logCFUse, ymax = logCFUmean+logCFUse), width = 0, size = 0.5)+
    scale_linetype_manual(values = c( "solid","dashed"), labels = c("P.aeruginosa","S. aureus"))+
    geom_hline(yintercept=8.88, color='gray', linetype = "dashed") +
    geom_hline(yintercept=8.98, color='gray', linetype = "solid") +
    labs(x = "Time (hours)", y = expression(paste(Log[10], " CFU")), linetype = "")+
    scale_y_continuous(limits = c(4,10))+
    scale_x_continuous(breaks = c(4,8,12,16))+
    theme_cowplot(12)
  legend <- get_legend(shakingMutPlot + theme(legend.box.margin = margin(0, 0, 0, -10)))
  prow <- plot_grid(shakingMutPlot+theme(legend.position="none"),staticMutPlot+theme(legend.position="none"),
                      align='vh',labels = c("Stirred", "Static"),hjust = c(-3, -3.8),nrow = 1)
    plot_grid(prow,legend,rel_widths = c(3, .4))
}

>>>>

```{r code Figure 1}
GraphGrowthCurveLog1<-function (rawData, statsData){
    shakingWtPlot<-ggplot()+
      geom_line(data=statsData%>% filter(condition == "Shaking" & strain == "wt"), aes(x=time, y=logCFUmean, linetype = bug), size=.5)+
      geom_errorbar(data=statsData%>% filter(condition == "Shaking" & strain == "wt"),aes(x = time, ymin = logCFUmean-logCFUse, ymax=logCFUmean+logCFUse), width = 0, size = 0.5)+
      geom_hline(yintercept=9.619, color='red', linetype = "solid") + #Pa stirred
      geom_hline(yintercept=9.13, color='red', linetype = "dashed") + #Sa stirred
      scale_linetype_manual(values = c( "solid","dashed"), labels = c(bquote(paste(italic('P. aeruginosa'))),bquote(paste(italic('S. aureus')))))+
      labs(x = "Time (hours)", y = expression(paste(Log[10], " CFU")), linetype = "Species")+
      scale_y_continuous(limits = c(4,10))+
      scale_x_continuous(breaks = c(4,8,12,16))+
      theme_cowplot(12)

    staticWtPlot<-ggplot()+
      geom_line(data=statsData%>% filter(condition == "Static" & strain == "wt"), aes(x=time, y=logCFUmean, linetype = bug), size=.5)+
      geom_errorbar(data=statsData%>% filter(condition == "Static" & strain == "wt"),aes(x = time,ymin=logCFUmean-logCFUse, ymax = logCFUmean+logCFUse), width = 0, size = 0.5)+
      scale_linetype_manual(values = c( "solid","dashed"), labels = c("stirred","static"))+
      geom_hline(yintercept=8.88, color='red', linetype = "dashed") +
      geom_hline(yintercept=9.07, color='red', linetype = "solid") +
      labs(x = "Time (hours)", y = expression(paste(Log[10], " CFU")), linetype = "")+
      scale_y_continuous(limits=c(4,10))+
      scale_x_continuous(breaks = c(4,8,12,16))+
      theme_cowplot(12)

    legend <- get_legend(shakingWtPlot + theme(legend.box.margin = margin(0, 0, 0, -10)))
    prow <- plot_grid(shakingWtPlot+theme(legend.position="none"),staticWtPlot+theme(legend.position="none"),
                      align='vh',labels = c("Mixed", "Static"),hjust = c(-3, -3.8),nrow = 1)
    plot_grid(prow,legend,rel_widths = c(3, .4))
    }
#Graph logs
GraphGrowthCurveLog1(PA14Data,PA14DataStats)
```
Figures 2 and 5 are on document : "Staph killing assays.Rmd"

```{r Figure 4}
GraphGrowthCurveLog2<-function (rawData, statsData){
  shakingMutPlot<-ggplot()+
    geom_line(data=statsData%>% filter(condition == "Shaking" & strain ==  "mut"), aes(x=time, y=logCFUmean, linetype = bug), size=.5)+
    geom_errorbar(data=statsData%>% filter(condition == "Shaking" & strain == "mut"), aes(x = time, ymin = logCFUmean-logCFUse, ymax = logCFUmean+logCFUse), width = 0, size = 0.5)+
    scale_linetype_manual(values = c( "solid","dashed"), labels = c(bquote(paste(italic('P. aeruginosa'))),bquote(paste(italic('S. aureus')))))+
    geom_hline(yintercept=9.13, color='red', linetype = "dashed") +
    geom_hline(yintercept=9.57, color='red', linetype = "solid") +
    labs(x = "Time (hours)", y = expression(paste(Log[10], " CFU")), linetype = "Species")+
    scale_y_continuous(limits = c(4,10))+
    scale_x_continuous(breaks = c(4,8,12,16))+
    theme_cowplot(12)

  staticMutPlot<-ggplot()+
    geom_line(data=statsData%>% filter(condition == "Static" & strain == "mut"), aes(x=time, y=logCFUmean, linetype = bug), size=.5)+
    geom_errorbar(data=statsData%>% filter(condition == "Static" & strain ==  "mut"), aes(x = time, ymin = logCFUmean-logCFUse, ymax = logCFUmean+logCFUse), width = 0, size = 0.5)+
    scale_linetype_manual(values = c( "solid","dashed"), labels = c(bquote(paste(italic('P. aeruginosa'))),bquote(paste(italic('S. aureus')))))+
    geom_hline(yintercept=8.88, color='red', linetype = "dashed") +
    geom_hline(yintercept=8.98, color='red', linetype = "solid") +
    labs(x = "Time (hours)", y = expression(paste(Log[10], " CFU")), linetype = "")+
    scale_y_continuous(limits = c(4,10))+
    scale_x_continuous(breaks = c(4,8,12,16))+
    theme_cowplot(12)
  legend <- get_legend(shakingMutPlot + theme(legend.box.margin = margin(0, 0, 0, -10)))
  prow <- plot_grid(shakingMutPlot+theme(legend.position="none"),staticMutPlot+theme(legend.position="none"),
                      align='vh',labels = c("Mixed", "Static"),hjust = c(-3, -3.8),nrow = 1)
    plot_grid(prow,legend,rel_widths = c(3, .4))
}

### 2020-07-28 Staph killing assays/Staph killing report.Rmd
updated figures

```{r zone of inhibition data input}
diskFile<-"/Users/jupabago/Documents/Whiteley/PROJECTS/pqsL SCFM2/Staph killing assays/disk difussion assays.csv"
diskData <-read.csv(diskFile)
diskData$strain <- factor(diskData$strain, levels = c("wt", "pqsl"))
diskData <- diskData %>% mutate(val1 = value1-6,val2 = value2-6,average = (val1+val2)/2 )#this removes the diameter of the disks
diskDataStats<-diskData %>% group_by(strain, condition, bio.rep) %>% summarise(meanTech = mean(average)) %>% ungroup() %>%group_by(strain, condition) %>% summarise(meanBio = mean(meanTech), SDBio = std.error(meanTech))
diskDataStats$strain <- factor(diskDataStats$strain, levels = c("wt", "pqsl"))
diskDataStats1<-diskDataStats %>% mutate(
  CondStrain = case_when(
    condition == "shaking" & strain == "pqsl" ~ "pqsLShake",
    condition == "shaking" & strain == "wt"~ "wtShake",
    condition == "static" & strain == "pqsl"~ "pqsLStat",
    condition == "static" & strain == "wt"~ "wtStat"))
```

@@

```{r final plots, version 1}
#this is the original Figure but it needs to be split in two parts
    DiskPlot<-ggplot()+
      geom_errorbar(data=diskDataStats1, aes(x = strain, ymin = meanBio - SDBio, ymax = meanBio + SDBio, group = condition),position = position_dodge(width = .6), width = .2)+
      geom_point(data=diskDataStats1, aes(x=strain, y=meanBio, shape =CondStrain),position = position_dodge2(width = .6), size=3)+
      scale_shape_manual(values = c(1,2,16,17),limits=c("wtStat","pqsLStat","wtShake", "pqsLShake"), 
                     labels = c(bquote(paste(italic('P. aeruginosa'), " wild-type static")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL static")),
                                bquote(paste(italic('P. aeruginosa'), " wild-type stirred")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL stirred"))))+
      scale_x_discrete(name ="", breaks = c("wt","pqsl"), labels =c("wild-type",expression(""*Delta*"pqsL")))+
      labs( y = "Zone of inhibition (mm)", shape = "Growth condtion") +
      theme_cowplot(12)

    luminescencePlot<-ggplot()+
      geom_line(data=luminescenceData1, aes(x=TimePoint, y=Mean, linetype = Condition, group = interaction (Strain, Condition)), size=.5,show.legend = FALSE)+
      geom_point(data=luminescenceData1, aes(x=TimePoint, y=Mean, shape = CondStrain), size=3)+
      scale_y_log10(name=expression(paste(italic('S. aureus'), " % luminescence vs control" )))+
      labs( x = "Time (minutes)",shape = "") +
      scale_shape_manual(values = c(1,2,16,17),limits=c("wtStat","pqsLStat","wtShake", "pqsLShake"), 
                     labels = c(bquote(paste(italic('P. aeruginosa'), " wild-type static")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL static")),
                                bquote(paste(italic('P. aeruginosa'), " wild-type stirred")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL stirred"))))+
      scale_linetype_manual(values =c("solid","dashed"))+
      theme(legend.position="top")+
      theme_cowplot(12)

    legendFigure2 <- get_legend(luminescencePlot + theme(legend.box.margin = margin(0, 0, 0, -10)))
    figure2 <- plot_grid(DiskPlot+theme(legend.position="none"),luminescencePlot+theme(legend.position="none"),
                      align='vh',labels = c("A", "B"),hjust = -1,nrow = 1)
    plot_grid(figure2,legendFigure2,rel_widths = c(6, 1.7))
```
```{r final plots, version 2}
#this figure needs to be split to show only wt data and mutant data and limit the time point to 7 minutes for the luminescence.

    DiskPlotWt<-ggplot()+
      geom_errorbar(data=diskDataStats1 %>% filter (strain == "wt"), aes(x = condition, ymin = meanBio - SDBio, ymax = meanBio + SDBio),width = .2)+
      geom_point(data=diskDataStats1 %>% filter (strain == "wt"), aes(x=condition, y=meanBio, shape =condition), size=3)+
      scale_shape_manual(values = c(1,2),limits=c("static","shaking"), 
                     labels = c(bquote(paste(italic('P. aeruginosa'), " wild-type static")),
                                bquote(paste(italic('P. aeruginosa'), " wild-type stirred"))))+
      scale_x_discrete(name ="", breaks = c("static","shaking"), labels =c("Static","Stirred"))+
    ylim(0,17.5)+
  labs( y = "Zone of inhibition (mm)", shape = "Growth condtion") +
      theme_cowplot(12)

    DiskPlotMut<-ggplot()+
      geom_errorbar(data=diskDataStats1 %>% filter (strain == "pqsl"), aes(x = condition, ymin = meanBio - SDBio, ymax = meanBio + SDBio),width = .2)+
      geom_point(data=diskDataStats1 %>% filter (strain == "pqsl"), aes(x=condition, y=meanBio, shape =condition), size=3)+
      scale_shape_manual(values = c(16,17),limits=c("static","shaking"), 
                     labels = c(bquote(paste(italic('P. aeruginosa '), Delta*"pqsL static")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL stirred"))))+
      ylim(0,17.5)+
      scale_x_discrete(name ="", breaks = c("static","shaking"), labels =c("Static","Stirred"))+
      labs( y = "Zone of inhibition (mm)", shape = "Growth condtion") +
      theme_cowplot(12)

    luminescencePlotWt<-ggplot()+
      geom_line(data=luminescenceData1 %>% filter (Strain == "wt") %>% filter (TimePoint<7), aes(x=TimePoint, y=Mean, group = Condition), size=.5, show.legend = FALSE)+
      geom_point(data=luminescenceData1%>% filter (Strain == "wt") %>% filter (TimePoint<7), aes(x=TimePoint, y=Mean, shape = Condition), size=3)+
      scale_y_log10(name=expression(paste(italic('S. aureus'), " % luminescence vs control" )), limits = c(1,121),labels = c(1, 10, 100),breaks = c(1, 10, 100))+
      labs( x = "Time (minutes)",shape = "") +
      scale_shape_manual(values = c(1,2),limits=c("static","shaking"), 
                     labels = c(bquote(paste(italic('P. aeruginosa'), " wild-type static")),
                                bquote(paste(italic('P. aeruginosa'), " wild-type stirred"))))+
      theme(legend.position="top")+
      theme_cowplot(12)

    luminescencePlotMut<-
      ggplot()+
      geom_line(data=luminescenceData1 %>% filter (Strain == "pqsL") %>% filter (TimePoint<8), aes(x=TimePoint, y=Mean, group = Condition), size=.5, show.legend = FALSE)+
      geom_point(data=luminescenceData1%>% filter (Strain == "pqsL") %>% filter (TimePoint<8), aes(x=TimePoint, y=Mean, shape = Condition), size=3)+
      scale_y_log10(name=expression(paste(italic('S. aureus'), " % luminescence vs control" )),limits = c(1,121), labels = c(1,10,100),breaks = c(1, 10, 100))+
      labs( x = "Time (minutes)",shape = "") +
      scale_shape_manual(values = c(16,17),limits=c("static", "shaking"), 
                     labels = c(bquote(paste(italic('P. aeruginosa '), Delta*"pqsL static")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL stirred"))))+
      theme(legend.position="top")+
      theme_cowplot(12)
#Figure 2

>>>>

```{r zone of inhibition data input}
diskFile<-"/Users/jupabago/Documents/Whiteley/PROJECTS/PaSa Project/Staph killing assays/disk difussion assays.csv"
diskData <-read.csv(diskFile)
diskData$strain <- factor(diskData$strain, levels = c("wt", "pqsl"))
diskData <- diskData %>% mutate(val1 = value1-6,val2 = value2-6,average = (val1+val2)/2 )#this removes the diameter of the disks
diskDataStats<-diskData %>% group_by(strain, condition, bio.rep) %>% summarise(meanTech = mean(average)) %>% ungroup() %>%group_by(strain, condition) %>% summarise(meanBio = mean(meanTech), SDBio = std.error(meanTech))
diskDataStats$strain <- factor(diskDataStats$strain, levels = c("wt", "pqsl"))
diskDataStats1<-diskDataStats %>% mutate(
  CondStrain = case_when(
    condition == "shaking" & strain == "pqsl" ~ "pqsLShake",
    condition == "shaking" & strain == "wt"~ "wtShake",
    condition == "static" & strain == "pqsl"~ "pqsLStat",
    condition == "static" & strain == "wt"~ "wtStat"))
```

@@

```{r final plots, version 1}
#this is the original Figure but it needs to be split in two parts
    DiskPlotSup<-ggplot()+
      geom_errorbar(data=diskDataStats1, aes(x = condition, ymin = meanBio - SDBio, ymax = meanBio + SDBio, group = strain),position = position_dodge(width = .6), width = .2)+
      geom_point(data=diskDataStats1, aes(x=condition, y=meanBio, shape =CondStrain, group = strain),position = position_dodge(width = .6), size=3)+
      scale_shape_manual(values = c(1,2,16,17),limits=c("wtStat","pqsLStat","wtShake", "pqsLShake"), 
                     labels = c(bquote(paste(italic('P. aeruginosa'), " wild-type static")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL static")),
                                bquote(paste(italic('P. aeruginosa'), " wild-type stirred")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL stirred"))))+
      scale_x_discrete(name ="", breaks = c("static","shaking"), labels =c("Static","Mixed"))+
      labs( y = "Zone of inhibition (mm)", shape = "Growth condtion") +
      theme_cowplot(12)

    luminescencePlotSup<-ggplot()+
      geom_line(data=luminescenceData1%>% filter (TimePoint<7), aes(x=TimePoint, y=Mean, linetype = Condition, group = interaction (Strain, Condition)), size=.5,show.legend = FALSE)+
      geom_point(data=luminescenceData1%>% filter (TimePoint<7), aes(x=TimePoint, y=Mean, shape = CondStrain), size=3)+
      scale_y_log10(name=expression(paste(italic('S. aureus'), " % luminescence vs control" )))+
      labs( x = "Time (minutes)",shape = "") +
      scale_shape_manual(values = c(1,2,16,17),limits=c("wtStat","pqsLStat","wtShake", "pqsLShake"), 
                     labels = c(bquote(paste(italic('P. aeruginosa'), " wild-type static")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL static")),
                                bquote(paste(italic('P. aeruginosa'), " wild-type stirred")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL stirred"))))+
      scale_linetype_manual(values =c("solid","dashed"))+
      theme(legend.position="top")+
      theme_cowplot(12)

    legendFigureS3 <- get_legend(luminescencePlotSup + theme(legend.box.margin = margin(0, 0, 0, -10)))
    figureS3 <- plot_grid(DiskPlotSup+theme(legend.position="none"),luminescencePlotSup+theme(legend.position="none"),
                      align='vh',labels = c("A", "B"),hjust = -1,nrow = 1)
    plot_grid(figureS3,legendFigureS3,rel_widths = c(6, 1.7))
```
```{r final plots, version 2}
#this figure needs to be split to show only wt data and mutant data and limit the time point to 7 minutes for the luminescence.

    DiskPlotWt<-ggplot()+
      geom_errorbar(data=diskDataStats1 %>% filter (strain == "wt"), aes(x = condition, ymin = meanBio - SDBio, ymax = meanBio + SDBio),width = .2)+
      geom_point(data=diskDataStats1 %>% filter (strain == "wt"), aes(x=condition, y=meanBio, shape =condition), size=3)+
      scale_shape_manual(values = c(1,2),limits=c("static","shaking"), 
                     labels = c(bquote(paste(italic('P. aeruginosa'), " wild-type static")),
                                bquote(paste(italic('P. aeruginosa'), " wild-type mixed"))))+
      scale_x_discrete(name ="", breaks = c("static","shaking"), labels =c("Static","Mixed"))+
    ylim(0,17.5)+
  labs( y = "Zone of inhibition (mm)", shape = "") +
      theme_cowplot(12)

    DiskPlotMut<-ggplot()+
      geom_errorbar(data=diskDataStats1 %>% filter (strain == "pqsl"), aes(x = condition, ymin = meanBio - SDBio, ymax = meanBio + SDBio),width = .2)+
      geom_point(data=diskDataStats1 %>% filter (strain == "pqsl"), aes(x=condition, y=meanBio, shape =condition), size=3)+
      scale_shape_manual(values = c(16,17),limits=c("static","shaking"), 
                     labels = c(bquote(paste(italic('P. aeruginosa '), Delta*"pqsL static")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL mixed"))))+
      ylim(0,17.5)+
      scale_x_discrete(name ="", breaks = c("static","shaking"), labels =c("Static","Mixed"))+
      labs( y = "Zone of inhibition (mm)", shape = "Growth condtion") +
      theme_cowplot(12)

    luminescencePlotWt<-ggplot()+
      geom_line(data=luminescenceData1 %>% filter (Strain == "wt") %>% filter (TimePoint<7), aes(x=TimePoint, y=Mean, group = Condition), size=.5, show.legend = FALSE)+
      geom_point(data=luminescenceData1%>% filter (Strain == "wt") %>% filter (TimePoint<7), aes(x=TimePoint, y=Mean, shape = Condition), size=3)+
      scale_y_log10(name=expression(paste(italic('S. aureus'), " % luminescence vs control" )), limits = c(1,121),labels = c(1, 10, 100),breaks = c(1, 10, 100))+
      labs( x = "Time (minutes)",shape = "") +
      scale_shape_manual(values = c(1,2),limits=c("static","shaking"), 
                     labels = c(bquote(paste(italic('P. aeruginosa'), " wild-type static")),
                                bquote(paste(italic('P. aeruginosa'), " wild-type mixed"))))+
      theme(legend.position="top")+
      theme_cowplot(12)

    luminescencePlotMut<-ggplot()+
      geom_line(data=luminescenceData1 %>% filter (Strain == "pqsL") %>% filter (TimePoint<7), aes(x=TimePoint, y=Mean, group = Condition), size=.5, show.legend = FALSE)+
      geom_point(data=luminescenceData1%>% filter (Strain == "pqsL") %>% filter (TimePoint<7), aes(x=TimePoint, y=Mean, shape = Condition), size=3)+
      scale_y_log10(name=expression(paste(italic('S. aureus'), " % luminescence vs control" )),limits = c(1,121), labels = c(1,10,100),breaks = c(1, 10, 100))+
      labs( x = "Time (minutes)",shape = "") +
      scale_shape_manual(values = c(16,17),limits=c("static", "shaking"), 
                     labels = c(bquote(paste(italic('P. aeruginosa '), Delta*"pqsL static")),
                                bquote(paste(italic('P. aeruginosa '), Delta*"pqsL mixed"))))+
      theme(legend.position="top")+
      theme_cowplot(12)
#Figure 2
